---
title: "portfolio"
format: html
editor: visual
author: Alondra Casas
---

### STAT 210 Portfolio


## Part 1 - Describe your Process

# Data Description

This is a dataset about shopping trends by demographic. I chose this dataset because I thought it would be interesting to look at shopping trends and purchases throughout seasons and potentially identify trends that cause the economy to bloom or crash. My data is from kaggle datasets and is called the Customer Shopping Trends dataset. The source describes the dataset as the journey into consumer insights and retail evolution with synthetic data. Variables in this dataset include items such as: gender, age, category, purchase amount and season.

# Data Visualization 
I want my final visualizations to consist of column plots and show proportion of people that buy specific clothing items. 

i want to highlight the average amount spent on shopping per season per season and idenity any distinct trends
I plan to answer these questions by making visualizations to show the proportion of categories such as gender and season.

I think a column of proportions would make the data easier to work with. 

You can add options to executable code like this

```{r}
#| echo: false
library("tidyverse")
library(ggplot2)
```


```{r}
star <- read_csv("data-raw/Casas Bat star Project Data - Sheet1.csv")
```



```{r}
#filtering for when larvae is at least 1 because I want to try running statistics and it doesn't like zeros, also might just make the graph look better
star_clean <- star |>
  filter(Larvae >= 1)

#making a column for status to label each container for each day with the exact counts of larvae and settled stars
star_clean2 <- star_clean |>
    pivot_longer(cols = c(Larvae, Settled), 
               names_to = "Status", 
               values_to = "Count")

#adding proportion to show a more distinct change in settlement rate
starclean3 <- star_clean |>
  mutate( Total = Larvae + Settled,
          propsettled = (Settled / Total)* 100)


```



```{r}
write_csv(star_clean, "data-clean/star_clean.csv")
write_csv(star_clean2, "data-clean/star_clean2.csv")
write_csv(starclean3, "data-clean/starclean3.csv")

```


```{r}
star |>
  ggplot(aes(x = Day, y = Settled, color = Treatment)) +
  geom_point(size = 0.5) +
  geom_smooth() +
  scale_x_continuous(breaks = 1:27) +
  labs(title = "Number of settled larvae over time", 
       x = "Days",
       y = "Number of settled larvae",
       color = "Treatment") +
  theme(legend.position = "bottom") 
```

